Contracts for unions and intersections are hard for Nickel

### 2021-05-26 paper.tex
More style conformance (authors, anonymous version)
Reading sections 1 & 2

\documentclass[sigplan,10pt,review]{acmart}
%\usepackage{amssymb}
\usepackage{mathtools,amsmath,amsfonts,amsxtra,amsthm,mathrsfs,mathpartir}
\usepackage{listings}
\usepackage{xcolor}
\usepackage{array}
\usepackage[toc,page]{appendix}
\usepackage{graphicx}
\usepackage{varioref}
\usepackage{verbatim,comment}
\usepackage{longtable}
\usepackage{tikz}
\usepackage{url}
\usepackage{hyperref}

@@

\title{Contracts for unions and intersections are hard for Nickel}
\author{Teodoro Freund, Yann Hamdaoui, Arnaud Spiwack}

@@

\section{Introduction}
Intro\unsure{todo}
\info{Goals of the paper: position paper: unions/intersections nice in
theory problematic in practice. In particular non-orthogonal with the
rest of the features of your language.}

@@

Enforcing pre- and post-conditions at runtime is a widely established practice.
In their foundational paper\cite{FindlerFelleisenHOContracts}, Findler and
Felleisen introduce a principled approach to run-time assertion checking that
nicely supports higher-order functions and introduces the notion of blame, which
is crucial to good error reporting. It became apparent later that their
contracts are closely related to the type casts introduced by gradual typing,
modulo blame: both \cite{FindlerMultiLang} and \cite{FelleisenInterLang} see the
value of contracts as a safe interface between typed and untyped code. In
\cite{WellTypedBlamed}, the authors precisely introduce a system integrating
gradual typing with contracts \textit{à la Findler \& Felleisen}.\unsure{The
historical bit should probably be moved to related works eventually} Nickel
adopts a similar type system, with both statically typed terms, dynamically
typed terms, and first-class contracts. In "first-class contracts", we include
the ability to write user-defined contracts. Because built-in types are too
simplistic for the purpose of data validation, the ability to write a small
validation function for, say, a port number, to have it immediately available as
a contract that can be composed with other type constructors\resolved{Needs code
to illustrate.}, is in our opinion one of the main appeal of Nickel's contracts
system. An example is given in Figure~\ref{fig:contract-composition}. The first
line defines a port contract from a predicate and use it in the definition of a
larger \lstinline+Configuration+ record contract. The \lstinline+Configuration+
contract checks that a value is a record with the following shape: it must have
a \lstinline+port+ field obeying the \lstinline+Port+ contract, and a
\lstinline+host+ field obeying a \lstinline+Host+ contract. It plays the role of
a \emph{schema}, used to validate the final data. The \lstinline+|+ operator
entails that a value respect a contract, akin to \lstinline+:+ for types. We
assume that a \lstinline+Host+ contract has been defined in a similar way.

@@

This type of contracts are present on many different languages,
for instance, the Eiffel programming language, the precursor
of the Design by Contract ideals, allows to assert
these kinds of expression as pre- and post-conditions on
functions\cite{EiffelDesignByContract}.
A somewhat newer language, the D programming
language, allows, among other things, to have invariants on classes,
written with flat contracts, that are
dinamically checked at the frontiers of every public method \cite{DLangContracts}.

@@

Enforcing pre- and post-conditions at runtime is a widely established practice.
In their foundational paper\cite{FindlerFelleisenHOContracts}, Findler and
Felleisen introduce a principled approach to run-time assertion checking that
nicely supports higher-order functions and introduces the notion of blame, which
is crucial to good error reporting. It became apparent later that their
contracts are closely related to the type casts introduced by gradual typing,
modulo blame: both \cite{FindlerMultiLang} and \cite{FelleisenInterLang} see the
value of contracts as a safe interface between typed and untyped code. In
\cite{WellTypedBlamed}, the authors precisely introduce a system integrating
gradual typing with contracts \textit{à la Findler \& Felleisen}.\unsure{The
historical bit should probably be moved to related works eventually} Nickel
adopts a similar type system, with both statically typed terms, dynamically
typed terms, and first-class contracts. In "first-class contracts", we include
the ability to write user-defined contracts. Because built-in types are too
simplistic for the purpose of data validation, the ability to write a small
validation function for, say, a port number, to have it immediately available as
a contract that can be composed with other type constructors\resolved{Needs code
to illustrate.}, is in our opinion one of the main appeal of Nickel's contracts
system. An example is given in Figure~\ref{fig:contract-composition}. The first
line defines a port contract from a predicate and use it in the definition of a
larger \lstinline+Configuration+ record contract. The \lstinline+Configuration+
contract checks that a value is a record with the following shape: it must have
a \lstinline+port+ field obeying the \lstinline+Port+ contract, and a
\lstinline+host+ field obeying a \lstinline+Host+ contract. It plays the role of
a \emph{schema}, used to validate the final data. The \lstinline+|+ operator
entails that a value respect a contract, akin to \lstinline+:+ for types. We
assume that a \lstinline+Host+ contract has been defined in a similar way.

>>>>

\documentclass[sigplan,10pt,review,anonymous]{acmart}
%\usepackage{amssymb}
\usepackage{mathtools,amsmath,amsfonts,amsxtra,amsthm,mathrsfs,mathpartir}
\usepackage{listings}
\usepackage{xcolor}
\usepackage{array}
\usepackage[toc,page]{appendix}
\usepackage{graphicx}
\usepackage{varioref}
\usepackage{verbatim,comment}
\usepackage{longtable}
\usepackage{tikz}
\usepackage{url}
\usepackage{hyperref}

@@

\title{Contracts for unions and intersections are hard for Nickel}
\author{Teodoro Freund}
\affiliation{
  \institution{???}
  \city{Buenos Aires}
  \country{Argentina}
}
\email{???}
\author{Yann Hamdaoui}
\affiliation{
  \institution{Tweag}
  \city{Paris}
  \country{France}
}
\email{yann.hamdaoui@tweag.io}
\author{Arnaud Spiwack}
\affiliation{
  \institution{Tweag}
  \city{Paris}
  \country{France}
}
\email{arnaud.spiwack@tweag.io}

@@

\unsure{TODO: CCS classification; Keywords}
\section{Introduction}
\info{Goals of the paper: position paper: unions/intersections nice in
theory problematic in practice. In particular non-orthogonal with the
rest of the features of your language.}

@@

Enforcing pre- and post-conditions at runtime is a widely established practice.
In their foundational paper\cite{FindlerFelleisenHOContracts}, Findler and
Felleisen introduce a principled approach to run-time assertion checking that
nicely supports higher-order functions and introduces the notion of blame, which
is crucial to good error reporting. It became apparent later that their
contracts are closely related to the type casts introduced by gradual typing,
modulo blame\unsure{What does ``modulo blame'' means here?}: both \cite{FindlerMultiLang} and \cite{FelleisenInterLang} see the
value of contracts as a safe interface between typed and untyped code. In
\cite{WellTypedBlamed}, the authors precisely introduce a system integrating
gradual typing with contracts \textit{à la Findler \& Felleisen}.\unsure{The
historical bit should probably be moved to related works eventually} Nickel
adopts a similar type system, with both statically typed terms, dynamically
typed terms, and first-class contracts. In "first-class contracts", we include
the ability to write user-defined contracts. Because built-in types are too
simplistic for the purpose of data validation, the ability to write a small
validation function for, say, a port number, to have it immediately available as
a contract that can be composed with other type constructors\resolved{Needs code
to illustrate.}, is in our opinion one of the main appeal of Nickel's contracts
system. An example is given in Figure~\ref{fig:contract-composition}. The first
line defines a port contract from a predicate and use it in the definition of a
larger \lstinline+Configuration+ record contract. The \lstinline+Configuration+
contract checks that a value is a record with the following shape: it must have
a \lstinline+port+ field obeying the \lstinline+Port+ contract, and a
\lstinline+host+ field obeying a \lstinline+Host+ contract. It plays the role of
a \emph{schema}, used to validate the final data. The \lstinline+|+ operator
entails that a value respect a contract, akin to \lstinline+:+ for types. We
assume that a \lstinline+Host+ contract has been defined in a similar
way.\unsure{What's missing in this section is an example of how
  contracts differ from standard dynamic type checking on some
  function. Namely, the quality of error messages.}

@@

This type of contracts are present on many different languages,
for instance, the Eiffel programming language\unsure{Citation?}, the precursor
of the Design by Contract ideals, allows to assert
these kinds of expression as pre- and post-conditions on
functions\cite{EiffelDesignByContract}.
A somewhat newer language, the D programming
language, allows, among other things, to have invariants on
classes\unsure{I (Arnaud) believe Eiffel has this feature as well,
  doesn't it?},
written with flat contracts, that are
dinamically checked at the frontiers of every public method \cite{DLangContracts}.

@@

Enforcing pre- and post-conditions at runtime is a widely established practice.
In their foundational paper\cite{FindlerFelleisenHOContracts}, Findler and
Felleisen introduce a principled approach to run-time assertion checking that
nicely supports higher-order functions and introduces the notion of blame, which
is crucial to good error reporting. It became apparent later that their
contracts are closely related to the type casts introduced by gradual typing,
modulo blame: both \cite{FindlerMultiLang} and \cite{FelleisenInterLang} see the
value of contracts as a safe interface between typed and untyped code. In
\cite{WellTypedBlamed}, the authors precisely introduce a system integrating
gradual typing with contracts \textit{à la Findler \& Felleisen}.\unsure{The
historical bit should probably be moved to related works eventually} Nickel
adopts a similar type system, with both statically typed terms, dynamically
typed terms, and first-class contracts. In "first-class contracts", we include
the ability to write user-defined contracts. Because built-in types are too
simplistic for the purpose of data validation, the ability to write a small
validation function for, say, a port number, to have it immediately available as
a contract that can be composed with other type constructors\resolved{Needs code
to illustrate.}, is in our opinion one of the main appeal of Nickel's contracts
system. An example is given in Figure~\ref{fig:contract-composition-appendix}. The first
line defines a port contract from a predicate and use it in the definition of a
larger \lstinline+Configuration+ record contract. The \lstinline+Configuration+
contract checks that a value is a record with the following shape: it must have
a \lstinline+port+ field obeying the \lstinline+Port+ contract, and a
\lstinline+host+ field obeying a \lstinline+Host+ contract. It plays the role of
a \emph{schema}, used to validate the final data. The \lstinline+|+ operator
entails that a value respect a contract, akin to \lstinline+:+ for types. We
assume that a \lstinline+Host+ contract has been defined in a similar way.

\begin{figure}
  \begin{center}
\begin{lstlisting}[language=nickel]
let Port = contracts.fromPred (fun value =>
  builtins.isNum value && value % 1 == 0
  && value >= 0 && value <= 65353) in
let Configuration = {
  host | #Host,
  port | #Port,
} in
{host = "nickel-lang.org", port = 80}
| #Configuration
\end{lstlisting}
  \end{center}
\caption{Contract composition}
\label{fig:contract-composition-appendix}
\medskip
\small
Defines a port contract from a predicate and use it in the definition of a
\lstinline+Configuration+ record contract. \lstinline+Configuration+ plays the
role of a \emph{schema} for the final data. The \lstinline+|+ operator applies a
contract to a value, similar to `:` for static types. We assume that a
\lstinline+Host+ contract has also been defined.
\end{figure}
